R package raincpc: Obtain and Analyze Rainfall data from the Climate Prediction Center (CPC)
========================================================

The Climate Prediction Center's (CPC, www.cpc.ncep.noaa.gov) daily rainfall data for the entire world, 1979 - present & 50-km resolution, is one of the few high quality and long-term observation-based rainfall products available for free. Data is available at CPC's (ftp site)[(ftp.cpc.ncep.noaa.gov/precip/CPC_UNI_PRCP/GAUGE_GLB/]. 

Some issues with size/format of the CPC data:
* there are too many daily files (365/366 per year * 34 years, separate folder for each year)
* file naming conventions have changed over time, one format prior to 2006 and couple of different formats afterwards
* downloading multiple files simultaneously from the CPC ftp site, using wget, does not seem to work properly
* file formats have changed over time - gzipped files prior to 2008 and binary files afterwards
* files are in binary format and there is no software/code readily available to process/plot the data

R package `raincpc`:
* Data for anytime period during 1979-2012 can be downloaded and processed
* Just two functions required: one to download the data (`cpc_get_rawdata`) and another to process the downloaded data (`cpc_read_rawdata`)
* Making spatial maps using the processed data is easy, via ggplot

Here are some examples.

Invoke required libraries.

```{r}
library(raincpc)

library(ggplot2)
library(reshape2)
```

Rainfall during the time of Hurricane Katrina, Aug - Sep of 2005.

```{r}
# obtain data
cpc_get_rawdata(2005, 8, 31, 2005, 8, 31) 
# process data
katrina <- cpc_read_rawdata(2005, 8, 31) 
```

re-orient data for plotting

```{r}
# function to flip a matrix upside down (change CPC orientation from S-N to N-S)
Fn_Flip_Matrix_Rows <- function(mat) {
  return (mat[nrow(mat):1,])
}
# function to rotate a matrix 90 degress clockwise for plotting only
# used to counteract the "image" function default behavior
Fn_Rotate_Matrix <- function(mat) {
  return (t(mat)[,nrow(mat):1])
}

katrina <- Fn_Rotate_Matrix(Fn_Flip_Matrix_Rows(katrina))

# plot data
gfx_map <- ggplot(data = melt(katrina)) 
gfx_map <- gfx_map + geom_raster(aes(Var1, Var2, fill = value)) 
gfx_map <- gfx_map + theme(axis.text = element_blank(), 
                           axis.ticks = element_blank())
gfx_map <- gfx_map + labs(x = NULL, y = NULL, fill = "Rain (mm/day)")
gfx_map <- gfx_map + ggtitle("Rainfall During Hurricane Katrina: 2005/8/31")
  
png("gfx_katrina.png", width = 800, height = 650)
plot(gfx_map)
garbage <- dev.off()
```

![figure 1] [fig1]

[fig1]: gfx_katrina.png "figure 1" 

