R package raincpc: Obtain and Analyze Rainfall data from the Climate Prediction Center (CPC)
========================================================

The Climate Prediction Center's (CPC, www.cpc.ncep.noaa.gov) daily rainfall data for the entire world, from 1979 to the present, at a resolution of 50 km (0.5 degrees lat-lon), is one of the few high quality and long-term observation-based rainfall products available for free. Data is available at CPC's (ftp site)[(ftp.cpc.ncep.noaa.gov/precip/CPC_UNI_PRCP/GAUGE_GLB/]. 

Some issues with size and format of the CPC data:
* there are too many daily files (365/366 per year * 34 years, separate folder for each year)
* file naming conventions have changed over time, one format prior to 2006 and couple of different formats afterwards
* downloading multiple files simultaneously from the CPC ftp site, using wget, does not seem to work properly
* file formats have changed over time, gzipped files prior to 2008 and binary files afterwards
* files are in binary format and there no software/code readily available to process/plot the data

R package `raincpc`:
* One function to download the data from CPC (`cpc_get_rawdata`) and one function to process the downloaded data (`cpc_read_rawdata`)
* Examples provided to plot spatial maps using ggplot



```{r}
summary(cars)
```

You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
plot(cars)
```

